---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- README.Rmd needs to be rendered to MD with devtools::build_readme() -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# edgemeta

<!-- badges: start -->
<!-- badges: end -->

The `edgemeta` package provides methods for frequentist random-effects meta-analysis using Edgington's method. It contains methods for estimation of the average effect and prediction of future effects, constructed using confidence distributions. 

The methods are implemented in the C++ programming language to ensure computational efficiency and are made available in the R environment using the Rcpp interface. 

## Estimation

A combined $p$-value function, or equivalently, confidence distribution, is constructed by combining one-sided $p$-value functions from individual studies. The confidence distribution of the average effect is then marginalized over the confidence distribution of the heterogeneity parameter to account for heterogeneity estimation uncertainty. Computation can be performed using a Monte Carlo algorithm generating independent samples, or using deterministic global adaptive quadrature integration. The latter is not recommended for few studies (3-5), since under these marginal distributions are slightly too wide. For mote than five studies, both methods typically produce very similar results.

## Prediction

The methods assumes that future effects and effects from observed studies are exchangeable and distributed according to a normal distribution. This joint effect distribution is maginalized over the confidence distribution from Edgington's method and over the confidence distribution of the heterogeneity parameter, thereby incorporating uncertainty about parameter estimation in both parameters. The predictive distributions are computed using a Monte Carlo sampling algorithm. Implemented are three forms of the predictive distribution:
- Full CD: Uses the full marginalization scheme and is therein the most accurate (recommended).
- Simplified CD: Uses a simplified approach, but accounts for uncertainty about both parameters.
- Fixed $\hat{\tau^2}$: Does not account for heterogeneity estimation uncertainty and is typically too narrow except for a large number of studies.

## Installation

You can install the `edgemeta` package from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("davidkronthaler-dk/edgemeta")
# or
# install.packages("devtools")
devtools::install_github("davidkronthaler-dk/edgemeta")
```

## Example

The `PredDist` function constructs an object of class `metaprediction`. Generic S3 methods (such as print, summary, or plot) are applicable. The function `PredDist` requires minimal input: a numeric vector of effect estimates, a corresponding vector of standard errors, and a string specifying the desired method. In case exact reproducibility of the result is required, a seed can be additionally set to ensure the latter.

Assume observed data (estimates and standard errors) from *k = 5* independent studies. The average effect can be estimated using the `remaeffect` function:

```{r example average effect}
library(edgemeta)

# Estimates and standard errors from 5 studies
es <- c(1.17,  2.20,  1.10, -0.0019, -1.33) 
se <- c(0.52, 0.93, 0.63, 0.3, 0.28)

# Average effect and confidence interval
me <- remaeffect(es = es, se = se, method = "MC") 

# or me <- remaeffect(es = es, se = se, method = "GAQ")
# or me <- remaeffect(es = es, se = se, method = "MHEU")

# Further information: 'help(remaeffect)'
```

A frequentist predictive distribution can be computed using:

```{r example predictive distribution}
# Computation of predictive distribution
pd <- PredDist(es = es, se = se, method = "FullCD")

# 95% prediction interval
pd$PI

# Plot the predictive distribution
plot(pd, param = "theta_new", breaks = 200, xlim = c(-7, 7))

# Futher information: 'help(PredDist)'
```

