---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- README.Rmd needs to be rendered to MD with devtools::build_readme() -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# edgemeta

<!-- badges: start -->
<!-- badges: end -->

The `edgemeta` package provides methods for frequentist random-effects meta-analysis using Edgington's method. It contains methods for estimation of the average effect and prediction of future study effects, constructed using confidence distributions. 

The methods are implemented in the C++ programming language to ensure computational efficiency and are made available in the R environment using the Rcpp interface. 

## Estimation

A combined $p$-value function, or equivalently, confidence distribution, is constructed by combining one-sided $p$-value functions from individual studies using Edgington's $p$-value combination method. The combined $p$-value function of the average effect is a confidence distribution of the average effect conditional on the heterogeneity parameter. To account for heterogeneity estimation uncertainty, it is then marginalized over the confidence distribution of the heterogeneity parameter. Computation can be performed using a Monte Carlo algorithm generating independent samples, or using deterministic global adaptive quadrature integration. The latter is not recommended for few studies (3-5), since under these marginal distributions are slightly too wide. For more than five studies, both methods typically produce very similar results.

## Prediction

The methods assumes that future effects and effects from observed studies are exchangeable and distributed according to a normal distribution. This joint effect distribution is maginalized over the confidence distribution from Edgington's method and over the confidence distribution of the heterogeneity parameter, thereby incorporating uncertainty about parameter estimation in both parameters. The predictive distributions are computed using a Monte Carlo sampling algorithm. Implemented are three forms of the predictive distribution:

- PCD-full: Uses the full marginalization scheme and is therein the most accurate (recommended).

- PCD-simplified: Uses a simplified approach, but accounts for uncertainty about both parameters.

- PCD-fixed: Does not account for heterogeneity estimation uncertainty and is typically too narrow except for a large number of studies.

## Installation

You can install the `edgemeta` package from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("davidkronthaler-dk/edgemeta")
# or
# install.packages("devtools")
devtools::install_github("davidkronthaler-dk/edgemeta")
```

## Example

Assume observed data (estimates and standard errors) from *k = 5* independent studies. The average effect can be estimated using the `remaeffect` function:

```{r remaeffect-example}
library(edgemeta)

# Estimates and standard errors from 5 studies
es <- c(1.17,  2.20,  1.10, -0.0019, -1.33) 
se <- c(0.52, 0.93, 0.63, 0.3, 0.28)

# Average effect and confidence interval
remaeffect(es = es, se = se, method = "MC") 

# or: remaeffect(es = es, se = se, method = "GAQ")

# Further information: 'help(remaeffect)'
```

A frequentist predictive distribution can be computed using:

```{r preddist-example}
# Computation of predictive distribution
pd <- PredDist(es = es, se = se, method = "PCD-full")

# or: PredDist(es = es, se = se, method = "PCD-simplified")
# or: PredDist(es = es, se = se, method = "PCD-fixed")

# 95% prediction interval
pd$PI

# Plot the predictive distribution
# plot(pd, param = "theta_new", breaks = 200, xlim = c(-7, 7))
# 
# Futher information: 'help(PredDist)'
```

